---
title: "Exercise 5"
output: github_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
install.packages("ggpubr")
```

## 0. Loading data and preparing variables

Here, I'm using the steps from exercise 2 example to create the necessary variables.

```{r prep-steps}
# set path for R to find our data
data_path <- "/Volumes/Study/MBA/PMBA/ORGB-690-051-People Analytics/Group project/Used data/"
library(arrow) # to be able to load data in the .parquet format

# read application data
app_data_sample <- read_parquet(paste0(data_path,"app_data_sample.parquet"))

library(gender)
#install_genderdata_package() # only run this line the first time you use the package, to get data for it

# get a list of first names without repetitions
examiner_names <- app_data_sample %>% 
  distinct(examiner_name_first)

examiner_names_gender <- examiner_names %>% 
  do(results = gender(.$examiner_name_first, method = "ssa")) %>% 
  unnest(cols = c(results), keep_empty = TRUE) %>% 
  select(
    examiner_name_first = name,
    gender,
    proportion_female
  )

# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
app_data_sample <- app_data_sample %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()

# Examiners' race
library(wru)

examiner_surnames <- app_data_sample %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()

examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_
  ))

examiner_race <- examiner_race %>% 
  select(surname,race)

app_data_sample <- app_data_sample %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()

# Examiner's tenure

library(lubridate) # to work with dates

examiner_dates <- app_data_sample %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date))) %>% 
  filter(year(end_date)<2018) %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    )

app_data_sample <- app_data_sample %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()


```
## Adding paygrade data

First, we load the paygrade file.

```{r load-paygrade-data}
examiner_gs <- read_csv(paste0(data_path,"examiner_gs.csv"))
examiner_ids <- read_csv(paste0(data_path,"examiner_ids.csv"))

```

### We need to replace various IDs with examiner ID

The ID fields in the `examiner_gs.csv` file don't match those in the application data file. Because we'll need to join these files later, we need to bring in the proper ID field, by using the cross-walk file `examiner_ids`.

```{r join-id}
examiner_gs <- examiner_gs %>% 
  left_join(examiner_ids) %>% 
  select(
    grade = examiner_grade,
    start_date,
    end_date,
    examiner_id = patex_id
  )
```

## Estimate time in grade

Now we need to estimate the average time each examiner spends in a given GS paygrade. Note that the less-biased way to do that is to exclude the latest or highest grade for each examiner. This is because after examiners reach the highest grade (which is usually grade 14 in this context), they cannot advance to the next grade. Imagine someone who advances through grades 11, 12 and 13 in just 18 months (so, the average for that examiner is 6 months per grade), but then stays in grade 14 and works at the agency for another ten years. If you were to average all grades, it would look like it took the examiner several years on average to get promoted through each grade. This is because the last grade biases the average positively. 

Similarly, for examiners who get promoted right before the end of our sample's time window, the very short time they stay in the highest observed grade will negatively bias the average. It will look like the examiner has progressed through the grades way faster than she did.

```{r get-time-in-grade}
time_in_grade <- examiner_gs %>% 
  mutate(
    start_date = mdy(start_date), # converting into proper date type
    end_date = mdy(end_date), # converting into proper date type
    days_in_grade = interval(start_date, end_date) %/% days(1)
  ) %>% 
  group_by(examiner_id) %>% 
  filter(grade!=max(grade, na.rm = TRUE)) %>% # dropping the highest grade record
  summarise(mean_days_in_grade = mean(days_in_grade, na.rm = TRUE))

time_in_grade
```

## Prepare application data

Let's get the measure of application processing time for each examiner. We'll do this by "collapsing" the dataframe from application level to examiner level (i.e., one record for each examiner, not multiple records).

```{r app-proc-time}
examiner_data <- app_data_sample %>% 
  filter(disposal_type!="PEND") %>% # here, we exclude in-process applications
  mutate(
    app_start_date = ymd(filing_date),
    app_end_date = case_when(
      disposal_type == "ISS" ~ ymd(patent_issue_date), # for issued patents
      disposal_type == "ABN" ~ ymd(abandon_date), # for abandoned applications
      TRUE ~ NA_Date_
    ),
    app_proc_days = interval(app_start_date, app_end_date) %/% days(1)) %>% 
  filter(app_proc_days>0 & app_proc_days < 3650) %>% # limit to 0-10 years
  group_by(examiner_id) %>% 
  summarise(
    app_count = n(),
    tc = min(tc, na.rm = TRUE),
    gender = first(gender),
    race = first(race),
    tenure_days = max(tenure_days, na.rm = TRUE),
    mean_app_proc_days = mean(app_proc_days, na.rm = TRUE)
  )

examiner_data
```

Now, let's join in the time in grade data.

```{r add-time-in-grade}
examiner_data <- examiner_data %>% 
  left_join(time_in_grade)

examiner_data
```

## Descriptive statistics and regressions

Let's run a couple of simple regressions.

```{r reg-sample}
library(modelsummary)
models <- list()
models[['m1']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days, data = examiner_data) 
models[['m2']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(gender), 
         data = examiner_data) 
modelsummary(models)
```
## Q1: How long does it take, on average, men and women to advance to the next pay grade? (Hint: you may want to exclude time in the highest paygrade for this analysis)
It took female agent 542 days to advance to the next pay grade, and took male agent 546 days to the next pat grade, which is very close to female. So we cannot observe obvious difference on days required for promotion based on gender. 


```{r reg-sample}
female_table <- examiner_data %>% 
  filter(gender == "female")
mean(female_table$mean_days_in_grade,na.rm=TRUE)

male_table <- examiner_data %>% 
  filter(gender == "male")
mean(male_table$mean_days_in_grade,na.rm=TRUE)
```

## Q2: Are these differences themselves different by examiners’ race?
It took white agent average 551 days to advance to the next pay grade, took Asian agent 534 days to the next pat grade, took black agent 589 days to the next pat grade, and took Hispanic agent 504 days to the next pat grade. The difference by race cannot be ignored, for example, the race with longest average days in one grade is about 17% more than the race with the shortest days in one grade. 

```{r reg-sample}
white_table <- examiner_data %>% 
  filter(race == "white")
mean(white_table$mean_days_in_grade,na.rm=TRUE)

asian_table <- examiner_data %>% 
  filter(race == "Asian")
mean(asian_table$mean_days_in_grade,na.rm=TRUE)

black_table <- examiner_data %>% 
  filter(race == "black")
mean(black_table$mean_days_in_grade,na.rm=TRUE)

hispanic_table <- examiner_data %>% 
  filter(race == "Hispanic")
mean(hispanic_table$mean_days_in_grade,na.rm=TRUE)
```
## Q3: Is there are relationship (an association) between average time in grade and average application prosecution time?
First, I calculated the correlation coefficient equals to 0.006 or 0.6%, which is very small meaning there is barely no relationship between average days in one grade and the average application prosecution time. 

Scatterplot also shows that there is no linear regression relationship between these two factors. 

By running a few regression models considering average application prosecution time, gender and race, I couldn't find obvious relationship between average time in one grade and the average application prosecution time. 
```{r reg-sample}

library(ggpubr)
```

```{r reg-sample}

cor(examiner_data$mean_days_in_grade,examiner_data$mean_app_proc_days, method = c("pearson", "kendall", "spearman"))
cor.test(examiner_data$mean_days_in_grade,examiner_data$mean_app_proc_days, method=c("pearson", "kendall", "spearman"))

ggscatter(examiner_data, x = "mean_days_in_grade", y = "mean_app_proc_days", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          )
names(examiner_data)

```

```{r reg-sample}
  
library(modelsummary)
models <- list()
models[['m1']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days, data = examiner_data) 
models[['m2']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(gender),data = examiner_data) 
models[['m3']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(race),data = examiner_data) 
models[['m4']] <- lm(mean_days_in_grade ~ 1 + mean_app_proc_days + as_factor(gender)+ as_factor(race), data = examiner_data) 
models[['m5']] <- lm(mean_days_in_grade ~ 1 + as_factor(gender) + as_factor(race), data = examiner_data)
modelsummary(models)
```

## Q4: Write a substantive conclusion, discussing what these results could mean. Make sure to discuss implications but also the limitations of your approach and possible threats to inference (this point must be done individually and not in a group)

My first conclusion is that we cannot use application processing time to predict the agents' promotion speed. The correlation between agents' promotion and race and gender is negligible too. It seems to me that, the application processing time cannot reflect agents' performance, and agents' promotion is not based how quickly they finish the applications. 

The second conclusion is that there is barely no competition exists in this organization, everyone can get a promotion after working for a similar period of time, which is slightly less than 2 years. 

The limitations of my analysis include: 
1. Maybe there are important factors that been used to evaluate agents' performance which do not been included in current data set. 
2. Maybe there are non-linear relationship exists between the average days in one grade and the factors been considered, which are not been checked here. 
